---
title: "HW3: Markdown file"
author: "Kevin Liu"
date: "2023-09-30"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem # 1

```{r}
library(p8105.datasets)
data("instacart")
instacart = 
  instacart |> 
  as_tibble()
```


##  Write a short description of the dataset, 

...noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

* The dataset instacart contains `r count(instacart)` observations and `r ncol(instacart)` variables.

* The dataset describes some form of online ordering system that keeps track of what is purchase in one order, how many, and what product as well as where the product is located by store department and aisle. 

* The dataset has many repeating rows for the same order and thus untidy, in the long form.

* Some key varaibles include the order_id (perhaps the designation for a certain shopper's order), product_id (what the shopper ordered), user_id (designation for the shopper), department (the store department where the item is located), and aisle (the store aisle where the item is located).

* For example, here is the order for shopper `r instacart$user_id[1]` with their order id `r instacart$order_id[1]`. 

```{r}
instacart |> 
  select(product_id, order_id, user_id, product_name, aisle, department) |> 
  head(8)
```

* There are `r instacart |> select(product_id) |> distinct() |> count()` products from `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

## Then, do or answer the following:

### How many aisles are there, and which aisles are the most items ordered from?

* There are `r instacart |> select(aisle) |>  distinct() |> count()` aisles.

The aisles that are the most ordered are:

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n)) |>
  head(3)
```

### Make a plot that...
shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### Make a table...
showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

### Make a table...
showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

# Problem #2