---
title: "HW3: Markdown file"
author: "Kevin Liu"
date: "2023-09-30"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem # 1

```{r}
library(p8105.datasets)
data("instacart")
instacart = 
  instacart |> 
  as_tibble()
```


##  Write a short description of the dataset, 

...noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

* The dataset instacart contains `r count(instacart)` observations and `r ncol(instacart)` variables.

* The dataset describes some form of online ordering system that keeps track of what is purchase in one order, how many, and what product as well as where the product is located by store department and aisle. 

* The dataset has many repeating rows for the same order and thus untidy, in the long form.

* Some key varaibles include the order_id (perhaps the designation for a certain shopper's order), product_id (what the shopper ordered), user_id (designation for the shopper), department (the store department where the item is located), and aisle (the store aisle where the item is located).

* For example, here is the order for shopper `r instacart$user_id[1]` with their order id `r instacart$order_id[1]`. 

```{r}
instacart |> 
  select(product_id, order_id, user_id, product_name, aisle, department) |> 
  head(8)
```

* There are `r instacart |> select(product_id) |> distinct() |> count()` products from `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

## Then, do or answer the following:

### How many aisles are there, and which aisles are the most items ordered from?

* There are `r instacart |> select(aisle) |>  distinct() |> count()` aisles.

The aisles that are the most ordered are:

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n)) |>
  head(3)
```

### Make a plot that...
shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### Make a table...
showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

### Make a table...
showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

# Problem #2

```{r}
#Import brfss data and convert into tibble dataframe
data("brfss_smart2010")
brfss_smart2010 = 
  brfss_smart2010 |> 
  as_tibble()
```

### __First__, do some data cleaning:

* format the data to use appropriate variable names;

```{r}
#COnvert variable names into snakecase
brfss_smart2010 =
  brfss_smart2010 |> 
  janitor::clean_names()
```

* focus on the “Overall Health” topic

```{r}
#Filter responses where topic == "OVerall Health Only"
brfss_overall_health =
  brfss_smart2010 |> 
  filter(topic == "Overall Health")

# brfss_overall_health |>
#   group_by(response) |> 
#   count()
```

* include only responses from “Excellent” to “Poor”:

```{r}
#Filter to include only responses that contain "Excellent", "Very good", "Good", "Fair", "Poor"
brfss_overall_health = 
  brfss_overall_health |> 
  filter(response == "Poor" | 
           response == "Fair" | 
           response == "Good" | 
           response == "Very good" | 
           response == "Excellent"
         )
```


* organize responses as a factor taking levels ordered from “Poor” to “Excellent”:

```{r}
#Convert variable type from char to factor
brfss_overall_health =
  brfss_overall_health |> 
  mutate(response = factor(
    response, 
    #we define the levels to have the 5 response types.
    levels = c ("Poor", "Fair", "Good", "Very good", "Excellent"), 
    #the categorical variable "response" is ordered so we indicate it as so
    ordered = TRUE)
    )
  
```

### Using this dataset, do or answer the following:

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
state20027cnt = 
  brfss_overall_health |> 
  filter(year == 2002) |>   #filter to include responses in 2002 only
  group_by(locationabbr) |>  #group by the state abbreviations
  summarize(num = n()) |>    #summarize the different response types 
  mutate(num = num/5) |>     #we know that there are 5 responses per location so we can get the number by                                   dividing
  arrange(desc(num))         #arrange by descending order for viewing ease

#output a table that shows stsates that have more than 7 locations and a state that had 6
state20027cnt |> 
  head(7) |> 
  knitr::kable(digits = 2)
```

* In 2002, PA, MA, NJ, CT, FL, and NC were states that were observed at 7 or more locations

```{r}
state20107cnt = 
  brfss_overall_health |> 
  filter(year == 2010) |>   #filter to include responses in 2010 only
  group_by(locationabbr) |>  #group by the state abbreviations
  summarize(num = n()) |>    #summarize the different response types 
  mutate(num = num/5) |>     #we know that there are 5 responses per location so we can get the number by                                   dividing
  arrange(desc(num))         #arrange by descending order for viewing ease

#output a table that shows stsates that have more than 7 locations and a state that had 6
state20107cnt |>
  head(15) |> 
  knitr::kable(digits = 2)
```

* In 2010, FL, NJ, TX, CA, MD, NC, NE, WA, MA, NY, OH, CO, PA, and SC, were states that were observed at 7 or more locations.

### Construct a dataset that is limited to...
Excellent responses, and contains, year, state, and a variable that averages the `data_value` across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_avg = 
  brfss_overall_health |> 
  select(year, locationabbr, response, data_value) |>
  filter(response == "Excellent") |> 
  group_by(year, locationabbr) |> 
  summarize(mean_data = mean(data_value, na.rm = TRUE))

brfss_avg |> 
  ggplot(aes(x = year, y = mean_data)) + 
  geom_line(aes(color = locationabbr), alpha = .5)
```

### Make a two-panel plot showing...
for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
#NY in 2006
brfssNY2006 =
  brfss_overall_health |> 
  filter(year == 2006) |> 
  filter(locationabbr == "NY")

#Plot
brfssNY2006plot = 
  brfssNY2006 |> 
  ggplot(aes(x = response, y = data_value, color = locationdesc)) +
  geom_point()

#NY in 2010
brfssNY2010 =
  brfss_overall_health |> 
  filter(year == 2010) |> 
  filter(locationabbr == "NY")

brfssNY2010plot =
  brfssNY2010 |> 
  ggplot(aes(x = response, y = data_value, color = locationdesc)) +
  geom_point()

brfssNY2006plot + brfssNY2010plot
```

